Follow-up Samples

### 2016-12-07 submission/manuscript_outline_20161024.Rmd
update abstract text

Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease.  Previous research suggests that the bacterial microbiome can be used as a biomarker for colorectal cancer (CRC).  These reports have mostly focused on investigating how the bacterial microbiome is used at a single point in time to predict disease.  In this study, we assessed whether a model built with bacterial microbiome data could accurately predict adenoma or carcinoma and adjust positive probability for lesion (adenoma or carcinoma) after surgical treatment.  This model was tested on a `r sum(initial_cm$ref_yes)` person group that included samples before and after treatment to allow for the assessment of how the model adjusts risk after treatment. The model used for prediction had an AUC of `r format(model_test_AUC_summary["full", "AUC"], digits = 3)`. For the follow up samples our Random Forest model significantly decreased the positive probability of lesion compared to the initial samples for both adenoma (P-value = `r format(IF_model_wilcox_paired_summary['all_adenoma', 'BH_correction'], scientific = TRUE, digits = 3)`) and carcinoma (P-value = `r format(IF_model_wilcox_paired_summary['carcinoma_only', 'BH_correction'], scientific = TRUE, digits = 3)`).  Our model predicted that `r format((followup_cm["pred_no", "ref_no"]/(sum(followup_cm$ref_no) + sum(followup_cm$ref_yes)))*100, digits = 3)`% of the 67-person cohort had normal colons and `r format(((followup_cm["pred_yes", "ref_no"] + sum(followup_cm$ref_yes))/(sum(followup_cm$ref_no) + sum(followup_cm$ref_yes)))*100, digits = 3)`% had a lesion.  Some OTUs that changed the most before and after treatment included OTUs that were affiliated with members of `r pvalue_common_imp_vars[order(pvalue_common_imp_vars$Pvalue), "Genus"][1:2]` and `r pvalue_common_imp_vars[order(pvalue_common_imp_vars$Pvalue), "Genus"][3]`. Our model suggests that treatment does significantly reduce the probability of having a colonic lesion.  Further surveillance of these individuals will enable us to determine whether models such as the one we present here can also be used to predict recurrence of colorectal cancer.

>>>>

Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease. Previous research suggests that the bacterial microbiome can be used as a biomarker for CRC.  In this study, we used a microbiome-based model that classified individuals as having an adenoma or carcinoma to assess whether the probability of having one of these lesions changed after surgical treatment.  We also characterized the change in the gut microbiota before and after the surgical treatment.  This model was tested on a `r sum(initial_cm$ref_yes)` person group that included samples before and after treatment to allow for the assessment of how the model adjusts risk after treatment. The model used for prediction had an AUC of `r format(model_test_AUC_summary["full", "AUC"], digits = 3)`. For the follow up samples our Random Forest model significantly decreased the positive probability of lesion compared to the initial samples for both adenoma (P-value = `r format(IF_model_wilcox_paired_summary['all_adenoma', 'BH_correction'], scientific = TRUE, digits = 3)`) and carcinoma (P-value = `r format(IF_model_wilcox_paired_summary['carcinoma_only', 'BH_correction'], scientific = TRUE, digits = 3)`).  Our model predicted that `r format((followup_cm["pred_no", "ref_no"]/(sum(followup_cm$ref_no) + sum(followup_cm$ref_yes)))*100, digits = 3)`% of the 67-person cohort had normal colons and `r format(((followup_cm["pred_yes", "ref_no"] + sum(followup_cm$ref_yes))/(sum(followup_cm$ref_no) + sum(followup_cm$ref_yes)))*100, digits = 3)`% had a lesion.  Some OTUs that changed the most before and after treatment included OTUs that were affiliated with members of `r pvalue_common_imp_vars[order(pvalue_common_imp_vars$Pvalue), "Genus"][1:2]` and `r pvalue_common_imp_vars[order(pvalue_common_imp_vars$Pvalue), "Genus"][3]`. Our model suggests that treatment does significantly reduce the probability of having a colonic lesion.  Further surveillance of these individuals will enable us to determine whether models such as the one we present here can also be used to predict recurrence of colorectal cancer.

### 2016-12-07 submission/manuscript_outline_20161024.tex
update abstract text

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests that the bacterial microbiome
can be used as a biomarker for colorectal cancer (CRC). These reports
have mostly focused on investigating how the bacterial microbiome is
used at a single point in time to predict disease. In this study, we
assessed whether a model built with bacterial microbiome data could
accurately predict adenoma or carcinoma and adjust positive probability
for lesion (adenoma or carcinoma) after surgical treatment. This model
was tested on a 66 person group that included samples before and after
treatment to allow for the assessment of how the model adjusts risk
after treatment. The model used for prediction had an AUC of 0.813. For
the follow up samples our Random Forest model significantly decreased
the positive probability of lesion compared to the initial samples for
both adenoma (P-value = 7.95e-08) and carcinoma (P-value = 7.95e-08).
Our model predicted that 36.4\% of the 67-person cohort had normal
colons and 63.6\% had a lesion. Some OTUs that changed the most before
and after treatment included OTUs that were affiliated with members of
Blautia, Clostridium\_XlVa and Escherichia/Shigella. Our model suggests
that treatment does significantly reduce the probability of having a
colonic lesion. Further surveillance of these individuals will enable us
to determine whether models such as the one we present here can also be
used to predict recurrence of colorectal cancer.

>>>>

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests that the bacterial microbiome
can be used as a biomarker for CRC. In this study, we used a
microbiome-based model that classified individuals as having an adenoma
or carcinoma to assess whether the probability of having one of these
lesions changed after surgical treatment. We also characterized the
change in the gut microbiota before and after the surgical treatment.
This model was tested on a 66 person group that included samples before
and after treatment to allow for the assessment of how the model adjusts
risk after treatment. The model used for prediction had an AUC of 0.813.
For the follow up samples our Random Forest model significantly
decreased the positive probability of lesion compared to the initial
samples for both adenoma (P-value = 7.95e-08) and carcinoma (P-value =
7.95e-08). Our model predicted that 36.4\% of the 67-person cohort had
normal colons and 63.6\% had a lesion. Some OTUs that changed the most
before and after treatment included OTUs that were affiliated with
members of Blautia, Clostridium\_XlVa and Escherichia/Shigella. Our
model suggests that treatment does significantly reduce the probability
of having a colonic lesion. Further surveillance of these individuals
will enable us to determine whether models such as the one we present
here can also be used to predict recurrence of colorectal cancer.
